Leaf traits drive differences in biomass partitioning among major plant functional types

### 2015-07-16 manuscript.Rnw
reformatted to 'normal' ms for NewPhyt or whatever
Newphyt bibtex style complete
not bold first para

\usepackage[sort&compress]{natbib}
\citestyle{nature}

@@

%--------------------------------------------------------------------------------------------%
%% Main text. Max words = 2000, first para = 200.

% First paragraph The first paragraph should be 180 to 200 words and should contain references. The paragraph should serve both as a general introduction to the topic and as a brief, non-technical summary of your main results and their implications. Ideally, the paragraph starts with two to three sentences of background explaining the topic and current understanding. This is followed by a phrase like “Here we present” to introduce the new work. In the second part, the main methods, findings, conclusions and implications should be described clearly and accessibly.

\section*{First paragraph}

\textbf{The partitioning of biomass into leaves and stems is one of the most uncertain  and influential components of  global vegetation models (GVMs), affecting the residence time and storage of carbon in ecosystems, and the response of net primary production to rising CO\textsubscript{2}\citep{dekauwe2014,friend2014,ise2010}. Although GVMs typically assume that the major woody plant functional types (PFTs) differ in biomass partitioning, empirical data have been lacking. Using a new global database on biomass partitioning from over 175 studies\citep{Falster2015}, we here show that partitioning between stems and leaves varies strongly between PFTs (as defined by deciduous vs. evergreen and gymnosperm vs. angiosperm). We show that major woody PFTs do not differ in the amount of leaf area supported, but rather the amount of leaf mass per unit above ground biomass. Moreover, we find that these differences relate to systematic shifts in leaf mass per area, a leading trait within the leaf economics spectrum \citep{wrightetal2004}. In addition, we show how a second trait, leaf area per stem basal area, explains additional variation in biomass partitioning among species; however this trait does not differ systematically among major woody PFTs. Combined, these results uncover fundamental rules in how plants are constructed and allow for systematic benchmarking of biomass partitioning routines in current GVMs.
}

\section*{Main text}

@@

@@

@@

\section*{Methods}

@@

% Figures A Nature Climate Change Letter can contain up to four display items (any combination of figures and/or tables). Figures should be sized in single or double column format. Multi-panel figures should be arranged so that the figure forms a neat rectangle.

>>>>

\usepackage{natbib}
\bibliographystyle{newphyt}

@@

\section*{Abstract}

The partitioning of biomass into leaves and stems is one of the most uncertain  and influential components of  global vegetation models (GVMs), affecting the residence time and storage of carbon in ecosystems, and the response of net primary production to rising CO\textsubscript{2}\citep{dekauwe2014,friend2014,ise2010}. Although GVMs typically assume that the major woody plant functional types (PFTs) differ in biomass partitioning, empirical data have been lacking. Using a new global database on biomass partitioning from over 175 studies\citep{Falster2015}, we here show that partitioning between stems and leaves varies strongly between PFTs (as defined by deciduous vs. evergreen and gymnosperm vs. angiosperm). We show that major woody PFTs do not differ in the amount of leaf area supported, but rather the amount of leaf mass per unit above ground biomass. Moreover, we find that these differences relate to systematic shifts in leaf mass per area, a leading trait within the leaf economics spectrum \citep{wrightetal2004}. In addition, we show how a second trait, leaf area per stem basal area, explains additional variation in biomass partitioning among species; however this trait does not differ systematically among major woody PFTs. Combined, these results uncover fundamental rules in how plants are constructed and allow for systematic benchmarking of biomass partitioning routines in current GVMs.

\section*{Introduction}

@@

\section*{Materials and Methods}

\section*{Results}

@@

\section*{Discussion}

@@

@@
